### 2017-10-09 main.tex
Different layout

\documentclass{article}

\usepackage{listings}

\title{beastscriptr: BEAUti for R}
\author{Rich\`el J.C. Bilderbeek}

\begin{document}

\maketitle

\abstract{
  \textbf{Motivation:} BEAST2 is a popular Bayesian phylogenetics tool.
  The program BEAUti allows for a user-friendly way to create the
  input files that BEAST2 needs. This is convenient to get started, yet
  does not facilitate the scripted use of BEAST2. 
  beastscriptr is an R package that can create BEAST2 input files from an R
  function call.\\
  \textbf{Results:} beastscriptr has only parts of the functionality of BEAUti, yet
  can be extended easily. beastcriptr, on the other hand, does allow for specifying
  to use phylogenies of fixed crown age.\\
  \textbf{Availability:} The package is free and is available from the official R package archive at 
  http://cran.r-project.org/src/contrib/PACKAGES.html\#beastscriptr. 
  beastscriptr is licensed under the GNU General Public License.
  \textbf{Key-words:} computational biology, evolution, phylogeny, BEAST2, R

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

BEAST2 \cite{bouckaert2014beast} is a popular Bayesian phylogenetics tool.
BEAST2 is bundled with the program called BEAUti \cite{drummond2012bayesian},
which  allows for a user-friendly way to create the
input files that BEAST2 needs. This is convenient to get started, yet
does not allow for the scripted use of BEAST2. beastscriptr is an R package 
that can create BEAST2 input files from an R function call.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Descriptions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Examples}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{beastscriptr development and other resources}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Citation of beastscriptr}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Scientists using beastscriptr in a published paper should cite this
article. Users can additionally cite the beastscriptr package 
directly. Citation information can be obtained by typing:

\begin{lstlisting}[language=R]
> citation("beastscriptr")
\end{lstlisting}

from within R.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We would like to thank the Center for Information Technology of the University of Groningen for their support
and for providing access to the Peregrine high performance computing cluster.

\appendix

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Approach}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The goal of the R package beastscriptr is to create the 
same BEAST2 input files that BEAUti creates, but from one R function call 
instead.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methods}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

By creating BEAST2 input files in BEAUti, 
the desired output of beastscriptr was determined. 
The functionality of BEAUti is added gradually (and is still growing).

beastscriptr has a novel functionality not built into BEAUti yet:
it allows for using phylogenies of a fixed crown age.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

beastscriptr does not support the full functionality of BEAUti. Considering
the size, age and number of plugins, this would be close to impossible.
To compensate for this, an extensible software architecture is used.

beastscriptr has minimal support for calling BEAST2 from within R and does
so for testing purposes.

beastscript has minimal support for parsing and interpreting BEAST2 output files,
for that the RBeast package is recommended.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

beastscriptr successfully creates valid BEAST2 input files. Unlike BEAUti,
beascriptr allows for setting a fixed crown age.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{natbib}
%\bibliographystyle{achemnat}
%\bibliographystyle{plainnat}
%\bibliographystyle{abbrv}
%\bibliographystyle{bioinformatics}
%
\bibliographystyle{plain}
%
\bibliography{document}

\begin{thebibliography}{}

\end{thebibliography}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}