### 2018-05-02 response_letter/response_letter.md
Implemented feedback reviewers

Dear Editor,

...

Kind regards,

Richèl J.C. Bilderbeek and Rampal S. Etienne