### 2019-05-02 figures/figures.tex
Figures are in 'figures.tex' (standalone) and 'figure_1.tex' (only the figure)

\documentclass[]{article}

% Bad idea
%\usepackage{fullpage}

%\documentclass[a0paper]{article}
% Must be at least 300 DPI

% No page numbering
\pagenumbering{gobble}

\usepackage{listings}
\usepackage{hyperref}
\usepackage{todonotes}
\usepackage{verbatim}
\usepackage{tikz}
\usepackage{tkz-graph}
\usepackage{pgf}
\usepackage{bm}

\usetikzlibrary{arrows,automata}
\usetikzlibrary{calc}
\usetikzlibrary{arrows.meta}

\begin{document}

\include{figure_1}

\end{document}